Estimation of cloud-corrected microwave humidity radiances with associated uncertainties

### 2020-11-18 Article/ici_submm_cloudcorr.tex
updated with Patrick's latest comments

The Ice Cloud Imager (ICI) is a new instrument on board EPS-SG satellite MetOp-SG (Meteorological Operational - Second Generation). MetOp-SG is scheduled for launch in 2023, and it will make ICI the first operational sensor observing Earth using sub-mm wavelengths. The main objective of ICI is to use high frequency channels for measuring ice cloud properties, and improve the representation of ice clouds in regional and global NWP models. ICI is a conically scanning radiometer that will measure 13 frequencies from 183\,GHz up to 664\, GHz.  Among all available channels, 183.31\,GHz, 325.15\,GHz and 448.00\,GHz, will measure vertical polarization;  while other channels around frequencies 243.20\,GHz and 664.00\,GHz are ``window channels'' and will measure both vertical and horizontal polarization. The instrument will observe earth from a mean altitude of 832\,km with sensor viewing angle 44.767$^\unit{^{\degree}}$(measured from nadir). For all the channels, the mean footprint size is about 15\,km, but the exact geo-location of samples differs. Therefore, a simultaneous utilization of data from different channels shall require remapping to a common footprint \citep{eriksson:towar:20}.

For this study, we conducted the forward simulations at frequencies: 183.31\,GHz, 325.15\,GHz, 448.00\,GHz and 664.00\,GHz (See Table~\ref{tab:ICI_MWI_channels}). For brevity, we assume that all simulations are mapped to a common footprint.

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The Arctic Weather Satellite (AWS) is a small satellite mission approved as European Space Agency (ESA)
Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels shall be included. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

In this study, forward simulations at frequencies 183.31\,GHz and 325.15\,GHz were conducted.

@@

ICI and AWS frequencies are simulated with Atmospheric Radiative Transfer Simulator (ARTS, \citet{buehler:artst:18}). CloudSat \citep{Stephens2002cloudsat} profiles during August 2015 are randomly selected for forward simulations. The input data are restricted between $60\unit{^{\degree}S}$ to $60\unit{^{\degree}N}$, and surface is below 500\,m. Both clear-sky and all-sky scenarios  are simulated. The complete simulation setup is described in Appendix \ref{appendix:ARTS_setup}. For ICI and AWS, 220\,000 and 143\,000 cases are simulated, respectively. All AWS sensor viewing angles (from $0^{\degree}$ to $45^{\degree}$) are simulated, but the results described in this study are based on nadir viewing angle. Simulations for all three sensors are noise free, so to incorporate the measurement uncertainties, whenever needed, gaussian noise is added to according to the channel NE$\Delta$T (see Table~\ref{tab:specifications_MWHS2} - Table~\ref{tab:specifications_AWS}).

@@

\subsection{Quantile regression neural networks}
\label{sec:QRNN}
%
The task that we aim to solve in this study is to predict the NFCS brightness
temperature \todo{$\ynfcs$} at a given 183\,GHz channel from a vector of
all-sky observations $\y$. Since the information-content of the
cloud-contaminated observations is certainly too low to solve this problem
exactly, a probabilistic formulation is appropriate here. The aim thus becomes
to predict the conditional distribution $p(\ynfcs | \y)$ of the NFCS brightness
temperatures $\ynfcs$ given the cloud-contaminated observations $\y$.

@@

QRNN's are trained to minimize the mean of the sum of the quantile loss functions
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions $\tau$, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data . The
quantile loss is used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution.

@@

The posterior distribution of $\ynfcs$ obtained from experiment 89+150\,GHz is similar to the one shown in Fig.~\ref{fig:posterior_distribution_I1V} and the distribution of point estimates is displayed in Fig.~\ref{fig:distribution_predicted_mwhs14}. The predicted values are able to correct most of the low brightness temperature cases, and overall a good match with the NFCS simulations is observed. However, QRNN is unable to predict the lowest clear-sky brightness temperatures, and cases with brightness temperatures around 260\,K occur frequently. The deviations of point estimates to NFCS simulations are shown in Fig.~\ref{fig:error_distribution_mwhs14}. The large negative deviations are removed (blue curve), but residual cloud impact is evident in the negative tail. Most of these residual cases have departures less than 10\,K. The appearance of a small positive tail also indicates overestimation in few cases. The corresponding error statistics(see Sect.~\ref{sec:validation}) are provided in  Table~\ref{tab:error_statistics_mwhs_14}. In the uncorrected all-sky simulations, the negative departures due to cloud impact lead to a large negative skewness(-12.72) and high bias (-0.84\,K). QRNN trained with 89 and 150\,GHz successfully corrects a major portion of the cloud affected cases, and the bias is reduced to -0.10\,K. However, negative skewness indicates presence of  uncorrected departures. Including 118\,GHz in the training gives comparable results as 89+150\,GHz. The error distribution has lower bias and are more symmetric, but the MAE and standard deviation remain unaltered. This indicates that the information from 118\,GHz channels can be beneficial in predicting few cases correctly, but the overall consequence is not exceptionally different from 89+150\,GHz. Similar is the case with using only 150\,GHz---the differences between the three experiments are negligible.

Further, we investigate whether filtering the predictions with low accuracy could help in improving the error distributions. For filtering such cases, we assume that predictions with cloud correction greater than 5\,K are associated with large deviations. In all three experiments, removing the cases with correction greater than 5\,K removes around 3\% of the data, but only a marginal positive impact on the accuracy is observed. The persistent negative skewness indicates presence of cases with residual cloud impact. Such cases are most likely to be associated with low cloud impact. Choosing a lower threshold can help in removing more partially corrected cases but at the cost of rejecting clear-sky cases. Since QRNN gives out NFCS values, choosing an unusually low correction threshold can also classify noisy clear-sky cases as cloudy. For example, for 89+150\,GHz, a threshold of 1.5\,K rejects almost 10\% of data, but the negative skewness is still not completely removed.

In spite of the fact that QRNN only provides a partial cloud correction, the results for  MWHS-2 channel 14 are comparable to what we achieve with existing cloud filtering techniques like scattering index (SI) and the filtering scheme by \citet{buehler:aclou:07}. SI uses the differences of brightness temperatures between 89 and 150\,GHz to identify cloud affected data, whilst in the study by \citet{buehler:aclou:07}, they use brightness temperature based threshold to filter out cloudy cases. They recommend a viewing angle dependent brightness temperature threshold at $183.31\pm1.00$ \,GHz and brightness temperature difference between $183.31\pm3.00$ \,GHz and $183.31\pm 1.00$ \,GHz as a measure of cloud impact. Further in the text we refer this filter as ``B183''. The results obtained with SI and B183 are displayed in last two columns of Table~\ref{tab:error_statistics_mwhs_14}. With SI threshold = 5\,K, almost 28\% of the data is rejected, yet the resulting error distributions are poorer as compared to QRNN. The low bias and skewness values indicate that most of the high negative departures are removed, but cases with low cloud impact pass the filter as clear. Similar is the case with B183. Here only 3\% of the data is filtered out but the overall statistics are worse than both QRNN and SI. The results from the two filters are not surprising as both are partial filters and aim at removing only cases with high ice content. The low hydrometeor impact cases remain unaltered.

The three experiments were also performed for other four 183\,GHz channels, and a similar performance was obtained (Table not shown). The positive effect of 118\,GHz was slightly higher for MWHS-2 channel 15, but for others, no notable effect was observed. In view of negligible performance differences between the three experiments, we consider the combination 89+159\,GHz to be the optimal. The results for other channels with this experiment are provided in Sect.~\ref{sec:mwhs_others}

@@

For channel 11, the error bias after correction is  -0.12\,K in comparison to 0.15\,K before correction. The decrease in bias is not significantly high but the strong negative tail  diminishes after correction, indicating removal of cases with large deviations. Nonetheless, the non-zero negative skewness also indicates presence of cases with residual cloud impact in the corrected dataset. Filtering the cases with high cloud impact has only a marginal positive effect. A similar performance is evident for channel 12. Correction reduces the bias to -0.12\,K from -0.29\,K and standard deviation to 1.08\,K from 2.02\.K. The MAE is approximately 16\% lower after correction, but still the negative skewness is not removed completely. For channel 13 and 15 also, a similar performance is seen. But in the latter, error distributions are most symmetric and have the largest spread as compared to other four channels. The effect of poor predictions is highest in channel 15 owing to its maximum sensitivity to hydrometeor impact.

@@

\subsection{Prediction uncertainty}
\label{sec:uncertainty_mwhs}
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/prediction_uncertainty_MWHS_14.pdf}	
	\caption{Examples of prediction uncertainties obtained from QRNN-single (89+150\,GHz) for MWHS-2 channel 14. 1500 randomly selected cases are shown. The quantiles have been plotted at equidistant points. The blue line represents a hypothetical Gaussian distribution of $270\pm1.0$\,K. For each quantile, the sample variation is also shown as box plot. }
	\label{fig:prediction_uncertainty_mwhs}	
\end{figure}
\begin{figure}[t]
	\includegraphics[height = 70mm]{Figures/calibration_QRNN_MWHS_14.pdf}	
	\caption{Calibration of the prediction intervals derived from QRNN-single (89+150\,GHz) for MWHS-2 channel 14, and prediction intervals derived from Gaussian error model for ECMWF NWP system (SI). For the former, the calibration for both the entire predicted dataset and the subset with correction greater than 5\,K are shown. }
	\label{fig:calibration_mwhs}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width=70mm]{Figures/deviation_posterior_mwhs_samples_14.pdf}	
	\caption{The distribution of predicted errors and observed errors  obtained from QRNN-single (89+150\,GHz) for MWHS-2 channel 14. The predicted errors are estimated as deviation of random samples from posteriori distribution to corresponding median values.}
	\label{fig:predicted_errors_mwhs}	
\end{figure}

@@

\subsection{Prediction uncertainty}
\label{sec:prediction_uncertainty}
%f 
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/prediction_uncertainty_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:prediction_uncertainty_mwhs}, but from QRNN-single for ICI channel I2V. The blue line represents a hypothetical Gaussian distribution of $270\pm0.65$\,K.}
	\label{fig:prediction_uncertainty_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[height = 70mm]{Figures/calibration_QRNN_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:calibration_mwhs}, but from QRNN-single for ICI channel I2V. }
	\label{fig:calibration_I1V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width=70mm]{Figures/deviation_posterior_samples_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:predicted_errors_mwhs}, but from QRNN-single for ICI channel I2V.}
	\label{fig:predicted_errors}	
\end{figure}

\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/cloud_impact_uncertainty_AWS.pdf}	
	\caption{The average confidence intervals (3$\sigma$) plotted against the magnitude of cloud impact for all AWS channels.}
	\label{fig:uncertainty_cloud_impact}	
\end{figure}
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{Figures/PDF_uncertainty_bins_QRNN-single.pdf}	
	\caption{Distribution of errors binned according to their uncertainty. Results are from QRNN-single for channels I1V, I2V and I3V.}
	\label{fig:error_distribution_uncertainty_bins}	
\end{figure*}
%f

@@

Figure~\ref{fig:calibration_I1V} shows the calibration of the prediction intervals for I2V. The predictions for the complete test dataset are well calibrated and follow the $y =x$ curve. On the other hand, when only the cases with  cloud correction greater than 10\,K are considered, the predictions in all the intervals are poorly calibrated. In this case, the predicted intervals are too wide over lower quantiles and too wide over higher quantiles.

@@

To conclude the results, we analyse if the uncertainty estimates given by QRNN are representative of prediction accuracy. Figure~\ref{fig:error_distribution_uncertainty_bins} shows the observed errors binned by their corresponding uncertainty in $\pm2\sigma$  confidence interval. For all three channels, the spread of error distribution increases as the uncertainty about the accuracy of the trial increases. The cases with high certainty have a narrow and sharp distribution, and the errors are mostly less than $\pm$2.5\,K. With increase in the uncertainty, frequency of cases with high accuracy decreases and the distributions spread out symmetrically to higher errors. Poor predictions occur more frequently when uncertainty is high. Cases with accurate predictions yet high uncertainty are also present. In spite of individual variations in the error distributions for each channel, predictions and their corresponding uncertainties follow the same relationship. Similar results are also obtained with AWS (not shown).

@@

Even though the cloud correction is partial, the performance is comparable or better to existing cloud filtering techniques like SI and scheme by \cite{buehler:aclou:07}. Note that both these techniques are a one for all approach for each 183\,GHz channel; thus if one observation is classified as cloudy by the filter, it is removed in all humidity channels. This increases the probability of erroneously removing clear observations. For high peaking channels of MWHS-2, both SI and QRNN give almost similar results, but with almost 28\% rejection rate. On the other hand, for low peaking channels, SI gives less accurate results than QRNN, as these channels have a stronger hydrometeor impact. This clearly indicates, that a channel specific approach like QRNN is more appropriate, and gives better performance.

The partial performance of QRNN is due to incomplete orthogonal information to 183\,GHz channels. The weighting functions of window channels 89 and 150\,GHz, and the two 118\,GHz channels peak in the lower troposphere. Among these four channels, 150\,GHz has the highest peaking function around 4\,km \citep{chen2020mwhs}. These channels can only provide coverage to the humidity channels in lower and mid troposphere; nonetheless, the 183\,GHz channels are sensitive to hydrometeor content up to 2\,km. Due to missing complementary information from other channels in the upper troposphere, QRNN fails at predicting these cases accurately. Such cases are mostly associated with thin cirrus clouds, which have very small influence at 89 and 150\,GHz. Without additional information from other channels we cannot expect QRNN to perform better. Among the other available channels, the overlapping weighting functions of 183\,GHz can provide auxiliary information to train QRNN. Results show that these channels indeed help in improving the training, yet the non-orthogonal information introduces highly correlated observational errors. The correlations for cloudy observations are not surprising as the cloud amount for different channels depends upon each other in a systematic way. However for clear-sky observations, the correlations should preferably be close to zero. This is observed to be true with QRNN-single, but QRNN-all fails at preserving the noise stochasticity. In the absence of hydrometeor impact, all 183\,GHz channels provide same information to the learning model, introducing redundancy. Redundant patterns in machine learning models often have undesirable effect on the predictive performance. In our application, redundant information does not affect the prediction accuracy, but introduces highly correlated observational errors between channels. Correlated errors are also undesirable for DA systems, but in future, if the operational centres progress with approaches dealing with correlated observation errors, for example as currently done for ATMS at ECMWF \citep{Weston2018eumATMS}, concurrent use of 89, 150 and 183\,GHz channels would give the best cloud correction performance.

>>>>

The Ice Cloud Imager (ICI) is a new instrument on board EPS-SG satellite MetOp-SG (Meteorological Operational - Second Generation). MetOp-SG is scheduled for launch in 2024, and it will make ICI the first operational sensor observing Earth using sub-mm wavelengths. The main objective of ICI is to use high frequency channels for measuring ice cloud properties, and improve the representation of ice clouds in regional and global NWP models. ICI is a conically scanning radiometer that will measure 13 frequencies from 183\,GHz up to 664\, GHz.  Among all available channels, 183.31\,GHz, 325.15\,GHz and 448.00\,GHz, will measure vertical polarization;  while other channels around frequencies 243.20\,GHz and 664.00\,GHz are ``window channels'' and will measure both vertical and horizontal polarization. The instrument will observe earth from a mean altitude of 832\,km with sensor viewing angle 44.767$^\unit{^{\degree}}$(measured from nadir). For all the channels, the mean footprint size is about 15\,km, but the exact geo-location of samples differs. Therefore, a simultaneous utilization of data from different channels shall require remapping to a common footprint \citep{eriksson:towar:20}.

For this study, we conducted the forward simulations of the channels around: 183.31\,GHz, 325.15\,GHz, 448.00\,GHz and 664.00\,GHz (Table~\ref{tab:ICI_MWI_channels}). For brevity, we assume that all simulations are mapped to a common footprint.

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The Arctic Weather Satellite (AWS) is a small satellite mission approved as European Space Agency (ESA)
Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels are presently planned. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

In this study, forward simulations of the channels around 183.31\,GHz and 325.15\,GHz were conducted.

@@

ICI and AWS frequencies are simulated with Atmospheric Radiative Transfer Simulator (ARTS, \citet{buehler:artst:18}). CloudSat \citep{Stephens2002cloudsat} profiles during August 2015 are randomly selected for forward simulations. The input data are restricted between $60\unit{^{\degree}S}$ to $60\unit{^{\degree}N}$, and surface is below 500\,m. Both clear-sky and all-sky scenarios  are simulated. The complete simulation setup is described in Appendix \ref{appendix:ARTS_setup}. For ICI and AWS, 220\,000 and 143\,000 cases are simulated, respectively. All AWS sensor viewing angles (from $0^{\degree}$ to $45^{\degree}$) are simulated, but the results described in this study are based on nadir viewing angle. Simulations for all three sensors are noise free, so to incorporate the measurement uncertainties, whenever needed, gaussian noise is added to according to the channel NE$\Delta$T (Table~\ref{tab:specifications_MWHS2} - Table~\ref{tab:specifications_AWS}).

@@

\subsection{Quantile regression neural networks}
\label{sec:QRNN}
%
The task that we aim to solve in this study is to predict the NFCS brightness
temperature $\ynfcs$ at a given 183\,GHz channel from a vector of
all-sky observations $\y$. Since the information-content of the
cloud-contaminated observations is certainly too low to solve this problem
exactly, a probabilistic formulation is appropriate here. The aim thus becomes
to predict the conditional distribution $p(\ynfcs | \y)$ of the NFCS brightness
temperatures $\ynfcs$ given the cloud-contaminated observations $\y$.

@@

QRNN's are trained to minimize the mean of the sum of the quantile loss functions
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions $\tau$, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data. The
quantile loss is also used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution. Note that CRPS is only used to evaluate hyper-parameter tuning.

@@

Posterior distributions of $\ynfcs$ obtained from experiment 89+150\,GHz are similar to the ones shown in Fig.~\ref{fig:posterior_distribution_I1V} and the distribution of point estimates is displayed in Fig.~\ref{fig:distribution_predicted_mwhs14}. The predicted values are able to correct most of the low brightness temperature cases, and overall a good match with the NFCS simulations is observed. However, QRNN is unable to predict the lowest clear-sky brightness temperatures, and cases with brightness temperatures around 260\,K occur too frequently. The deviations of point estimates to NFCS simulations are shown in Fig.~\ref{fig:error_distribution_mwhs14}. The large negative deviations are removed (blue curve), but residual cloud impact is evident in the negative tail. Most of these residual cases have departures less than 10\,K. The appearance of a small positive tail also indicates overestimation in few cases. The corresponding error statistics (see Sect.~\ref{sec:validation}) are provided in  Table~\ref{tab:error_statistics_mwhs_14}. In the uncorrected all-sky simulations, the negative departures due to cloud impact lead to a large negative skewness(-12.72) and high bias (-0.84\,K). QRNN trained with 89 and 150\,GHz successfully corrects a major portion of the cloud affected cases, and the bias is reduced to -0.10\,K. However, negative skewness indicates presence of  uncorrected departures. Including 118\,GHz in the training gives a small improvement compared to 89+150 GHz alone. The error distribution has lower bias and are more symmetric, but the MAE and standard deviation remain unaltered. This indicates that the information from 118\,GHz channels can be beneficial in predicting few cases correctly, but the overall performance is not exceptionally different from 89+150\,GHz. Similar is the case with using only 150\,GHz---the differences between the three experiments are negligible.

Further, we investigate whether filtering the predictions with low accuracy could help in improving the error distributions. For filtering such cases, we assume that predictions with cloud correction greater than 5\,K are associated with large deviations. In all three experiments, removing the cases with correction greater than 5\,K removes around 3\% of the data, but only a marginal positive impact on the accuracy is observed. The persistent negative skewness, even after filtering, indicates presence of cases with residual cloud impact. Such cases are most likely to be associated with low or medium cloud impact. Choosing a lower threshold can help in removing more partially corrected cases but at the cost of rejecting clear-sky cases. Since QRNN gives out NFCS values, choosing an unusually low correction threshold can also classify noisy clear-sky cases as cloudy. For example, for 89+150\,GHz, a threshold of 1.5\,K rejects almost 10\% of data, but the negative skewness is still not completely removed.

In spite of the fact that QRNN only provides a partial cloud correction, the results for  MWHS-2 channel 14 are better than what we achieve with existing cloud filtering techniques like scattering index (SI) and the filtering scheme by \citet{buehler:aclou:07}. SI uses the differences of brightness temperatures between 89 and 150\,GHz to identify cloud affected data, whilst in the study by \citet{buehler:aclou:07}, they use brightness temperature based threshold to filter out cloudy cases. They recommend a viewing angle dependent brightness temperature threshold at $183.31\pm1.00$ \,GHz and brightness temperature difference between $183.31\pm3.00$ \,GHz and $183.31\pm 1.00$ \,GHz as a measure of cloud impact. Further in the text we refer this filter as ``B183''. The results obtained with SI and B183 are displayed in last two columns of Table~\ref{tab:error_statistics_mwhs_14}. With SI threshold = 5\,K, almost 28\% of the data is rejected, yet the resulting error distributions are poorer as compared to QRNN. The low bias and skewness values indicate that most of the high negative departures are removed, but cases with low cloud impact pass the filter as clear. Similar is the case with B183. Here only 3\% of the data is filtered out but the overall statistics are worse than both QRNN and SI. The results from the two filters are not surprising as both are partial filters and aim at removing only cases with high ice content. The low hydrometeor impact cases remain unaltered.

The three experiments were also performed for other four 183\,GHz channels, and a similar performance was obtained (not shown). The positive effect of 118\,GHz was slightly higher for MWHS-2 channel 15, but for others, no notable effect was observed. In view of negligible performance differences between the three experiments, we consider the combination 89+159\,GHz to be the optimal. The results for other channels with this experiment are provided in Sect.~\ref{sec:mwhs_others}

@@

For channel 11, the error bias after correction is  -0.12\,K in comparison to 0.15\,K before correction. The decrease in bias is not significantly high but the strong negative tail  diminishes after correction, indicating removal of cases with large deviations. Nonetheless, the non-zero negative skewness also indicates presence of cases with residual cloud impact in the corrected dataset. Filtering the cases with high cloud impact has only a marginal positive effect. A similar performance is evident for channel 12. Correction reduces the bias to -0.12\,K from -0.29\,K and standard deviation to 1.08\,K from 2.02\.K. The MAE is approximately 16\% lower after correction, but still the negative skewness is not removed completely. For channel 13 and 15 also, a similar performance is seen. But in the latter, error distributions are more symmetric and have the largest spread as compared to other four channels. The effect of poor predictions is highest in channel 15 owing to its maximum sensitivity to hydrometeor impact.

@@

\subsection{Prediction uncertainty}
\label{sec:uncertainty_mwhs}
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/prediction_uncertainty_MWHS_14.pdf}	
	\caption{Examples of prediction uncertainties obtained from QRNN-single (89+150\,GHz) for MWHS-2 channel 14. 1500 randomly selected cases are shown. The quantiles have been plotted at equidistant points. The blue line represents a Gaussian distribution with a standard deviation of $1.0$\,K. For each quantile, the sample variation is also shown as box plot. }
	\label{fig:prediction_uncertainty_mwhs}	
\end{figure}
\begin{figure}[t]
	\includegraphics[height = 70mm]{Figures/calibration_QRNN_MWHS_14.pdf}	
	\caption{Calibration of the prediction intervals derived from QRNN-single (89+150\,GHz) for MWHS-2 channel 14, and prediction intervals derived from Gaussian error model for ECMWF NWP system (SI). For the former, the calibration for both the entire predicted dataset and the subset with correction greater than 5\,K are shown. }
	\label{fig:calibration_mwhs}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width=70mm]{Figures/deviation_posterior_mwhs_samples_14.pdf}	
	\caption{The distribution of predicted errors and observed errors  obtained from QRNN-single (89+150\,GHz) for MWHS-2 channel 14. The predicted errors are estimated as deviation of random samples from posteriori distribution to corresponding median values.}
	\label{fig:predicted_errors_mwhs}	
\end{figure}

@@

\subsection{Prediction uncertainty}
\label{sec:prediction_uncertainty}
%f 
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/prediction_uncertainty_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:prediction_uncertainty_mwhs}, but from QRNN-single for ICI channel I2V. The blue line represents a Gaussian distribution with a standard deviation of $0.65$\,K.}
	\label{fig:prediction_uncertainty_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[height = 70mm]{Figures/calibration_QRNN_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:calibration_mwhs}, but from QRNN-single for ICI channel I2V. }
	\label{fig:calibration_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width=70mm]{Figures/deviation_posterior_samples_I2V.pdf}	
	\caption{Same as Fig.~\ref{fig:predicted_errors_mwhs}, but from QRNN-single for ICI channel I2V.}
	\label{fig:predicted_errors}	
\end{figure}

\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/cloud_impact_uncertainty_AWS.pdf}	
	\caption{The average confidence intervals ($\pm3\sigma$) plotted against the magnitude of cloud impact for all AWS channels.}
	\label{fig:uncertainty_cloud_impact}	
\end{figure}
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{Figures/PDF_uncertainty_bins_QRNN-single.pdf}	
	\caption{Distribution of errors binned according to their uncertainty. Results are from QRNN-single for channels I1V, I2V and I3V.}
	\label{fig:error_distribution_uncertainty_bins}	
\end{figure*}
%f

@@

Figure~\ref{fig:calibration_I2V} shows the calibration of the prediction intervals for I2V. The predictions for the complete test dataset are well calibrated and follow the $y =x$ curve. Similar is the case, when cases with correction greater than 5\,K are considered. On the other hand, when  the cases with cloud correction greater than 10\,K are considered, the calibration is only slightly worse. In spite of the fact that such cases are few (~2\%), the high calibration indicates that QRNN is also successful in predicting the uncertainties associated with rare cases. For other two channels also the predictions are well calibrated, except for cases with correction greater than 5\,K in channel I1V (not shown).

@@

To conclude the results, we analyse if the uncertainty estimates given by QRNN are representative of prediction accuracy. Figure~\ref{fig:error_distribution_uncertainty_bins} shows the observed ICI errors binned by their corresponding uncertainty in $\pm2\sigma$  confidence interval. For all three channels, the spread of error distribution increases as the uncertainty about the accuracy of the trial increases. The cases with high certainty have a narrow and sharp distribution, and the errors are mostly less than $\pm$2.5\,K. With increase in the uncertainty, frequency of cases with high accuracy decreases and the distributions spread out symmetrically to higher errors. Poor predictions occur more frequently when uncertainty is high. Cases with accurate predictions yet high uncertainty are also present. In spite of individual variations in the error distributions for each channel, predictions and their corresponding uncertainties follow the same relationship. Similar results are also obtained with AWS (not shown).

@@

Even though the cloud correction is partial, the performance is comparable or better to existing cloud filtering techniques like SI and scheme by \cite{buehler:aclou:07}. Note that both these techniques are a one for all approach for each 183\,GHz channel; thus if one observation is classified as cloudy by the filter, it is removed in all humidity channels. This increases the probability of erroneously removing clear observations. For high peaking channels of MWHS-2, both SI and QRNN give almost similar results, but with almost 28\% rejection rate in the former. On the other hand, for low peaking channels, SI gives less accurate results than QRNN, as these channels have a stronger hydrometeor impact. This clearly indicates, that a channel specific approach like QRNN is more appropriate, and gives better performance.

The partial performance of QRNN is due to incomplete orthogonal information to 183\,GHz channels. The weighting functions of window channels 89 and 150\,GHz, and the two 118\,GHz channels peak in the lower troposphere. Among these four channels, 150\,GHz has the highest peaking function around 4\,km \citep{chen2020mwhs}. These channels can only provide coverage to the humidity channels in lower and mid troposphere; nonetheless, the 183\,GHz channels are sensitive to hydrometeor content up to 10\,km. Due to missing complementary information from other channels in the upper troposphere, QRNN fails at predicting these cases accurately. Such cases are mostly associated with thin cirrus clouds, which have very small influence at 89 and 150\,GHz. Without additional information from other channels we cannot expect QRNN to perform better. Among the other available channels, the overlapping weighting functions of 183\,GHz can provide auxiliary information to train QRNN. Results show that these channels indeed help in improving the training, yet the non-orthogonal information introduces highly correlated observational errors. The correlations for cloudy observations are not surprising as the cloud amount for different channels depends upon each other in a systematic way. However for clear-sky observations, the correlations should preferably be close to zero. This is observed to be true with QRNN-single, but QRNN-all fails at preserving the noise stochasticity. In the absence of hydrometeor impact, all 183\,GHz channels provide same information to the learning model, introducing redundancy. Redundant patterns in machine learning models often have undesirable effect on the predictive performance. In our application, redundant information does not affect the prediction accuracy, but introduces highly correlated observational errors between channels. Correlated errors are also undesirable for DA systems, but in future, if the operational centres progress with approaches dealing with correlated observation errors, for example as currently done for ATMS at ECMWF \citep{Weston2018eumATMS}, concurrent use of 89, 150 and 183\,GHz channels would give the best cloud correction performance.
